% Copyright (c) 2005 Nokia Corporation
%
% Licensed under the Apache License, Version 2.0 (the "License");
% you may not use this file except in compliance with the License.
% You may obtain a copy of the License at
%
%     http://www.apache.org/licenses/LICENSE-2.0
%
% Unless required by applicable law or agreed to in writing, software
% distributed under the License is distributed on an "AS IS" BASIS,
% WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
% See the License for the specific language governing permissions and
% limitations under the License.
\chapter{Terms and Abbreviations}

\label{sec:terms}
The following list defines the terms and abbreviations used in this 
document:
\begin{longtableii}{p{1in}|p{4.5in}}{textrm}{Term}{Definition}
\lineii{AAC; Adaptive Audio Coding}{AAC provides basically the same sound quality as MP3 while using a smaller bit rate. AAC is mainly used to compress music.}
\lineii{Advertise}{Advertise service in Bluetooth makes it known that a certain Bluetooth service is available. }
\lineii{AMR}{Adaptive Multi-rate Codec file format.}
\lineii{API }{Application Programming Interface}
\lineii{Bluetooth }{Bluetooth is a technology for wireless communication between devices that is based on a low-cost short-range radio link.}
\lineii{BPP }{Bits Per Pixel }
\lineii{C STDLIB}{Symbian OS's implementation of the C standard library}
\lineii{Dialog}{A temporary user interface window for presenting context-specific information to the user, or prompting for information in a specific context.}
\lineii{Discovery}{Discovery is a process where Bluetooth finds other nearby Bluetooth devices and their advertised services.}
\lineii{DLL }{Dynamic link library}
\lineii{GSM; Global System for Mobile communication}{GSM is a digital mobile telephone system that uses a variation of time division multiple access. It digitizes and compresses data, then sends it down a channel with two other streams of user data, each in its own time slot.}
\lineii{GUI}{Graphical User Interface}
\lineii{I/O }{input/output}
\lineii{IP }{Internet Protocol}
\lineii{MBM; MultiBitMap}{The native Symbian OS format used for pictures. MBM files can be generated with the \code{bmconv.exe} tool included in the S60 SDK.}
\lineii{MIDI; Musical Instrument Digital Interface}{A protocol and a set of commands for storing and transmitting information about music.}
\lineii{MIF; Multi-Image File}{MIF files are similar to MBM files and can contain compressed SVG-T files. This file type can be generated with the \code{MifConv.exe} tool.}
\lineii{MIME; Multipurpose Internet Mail Extensions}{MIME is an extension of the original Internet e-mail protocol that can be used to exchange different kinds of data files on the Internet.}
\lineii{MP3}{A standard technology and format for compressing a sound sequence into a very small file while preserving the original level of sound quality when it is played.}
\lineii{OS }{Operating System}
\lineii{Real Audio}{An audio format developed by Real Networks.}
\lineii{RDBMS}{Relational database management system}
\lineii{SMS; Short Message System (within GSM)}{SMS is a service for sending messages of up to 160 characters, or 224 characters if using a 5-bit mode, to mobile phones that use GSM communication.}
\lineii{Softkey}{Softkey is a key that does not have a fixed function nor a function label printed on it. On a phone, selection keys reside below or above on the side of the screen, and derive their meaning from what is presently on the screen.}
\lineii{SQL }{Structured Query Language}
\lineii{SVG, SVG-T; Scalable Vector Graphics (-Tiny)}{XML-based vector graphics format for describing two-dimensional graphics and graphical applications.}
\lineii{Twip}{Twips are screen-independent units to ensure that the proportion of screen elements are the same on all display systems. A twip is defined as 1/1440 of an inch, or 1/567 of a centimeter.}
\lineii{UI}{User Interface}
\lineii{UI control}{UI control is a GUI component that enables user interaction and represents properties or operations of an object.}
\lineii{WAV }{A file format for recording sound, especially in multimedia applications. }
\end{longtableii}
